
> I've been using keras and TensorFlow for a while now - and love its simplicity and straight-forward way to modeling. As part of the latest update to my [Workshop about deep learning with R and keras](https://shirinsplayground.netlify.com/2018/05/deep_learning_keras_tensorflow_18_07/) I've added a new example analysis:

https://shirinsplayground.netlify.com/2018/06/keras_fruits/

```{r}
# load libraries
library(tidyverse)
library(keras)
```

```{r}
# check if keras is available
is_keras_available()
```



### Loading images

The handy `image_data_generator()` and `flow_images_from_directory()` functions can be used to load images from a directory. If you want to use data augmentation, you can directly define how and in what way you want to augment your images with `image_data_generator`. Here I am not augmenting the data, I only scale the pixel values to fall between 0 and 1.

```{r}
train_data_gen <- image_data_generator(
  rescale = 1/255,
  validation_split = 0.3
)
```

```{r}
# training images
train_image_array_gen <- flow_images_from_directory(train_image_files_path, 
                                          train_data_gen,
                                          subset = 'training',
                                          target_size = target_size,
                                          class_mode = "categorical",
                                          classes = fruit_list,
                                          seed = 42)

# validation images
valid_image_array_gen <- flow_images_from_directory(train_image_files_path, 
                                          train_data_gen,
                                          subset = 'validation',
                                          target_size = target_size,
                                          class_mode = "categorical",
                                          classes = fruit_list,
                                          seed = 42)
```

```{r}
cat("Number of images per class:")
table(factor(train_image_array_gen$classes))

cat("\nClass label vs index mapping:\n")
train_image_array_gen$class_indices
```

### Define model

```{r}
# number of training samples
train_samples <- train_image_array_gen$n
# number of validation samples
valid_samples <- valid_image_array_gen$n

# define batch size and number of epochs
batch_size <- 32
epochs <- 10
```

```{r warning=FALSE, message=FALSE}
# initialise model
model <- keras_model_sequential()

# add layers
model %>%
  layer_conv_2d(filter = 32, kernel_size = c(3,3), padding = "same", input_shape = c(img_width, img_height, channels)) %>%
  layer_activation("relu") %>%
  
  # Second hidden layer
  layer_conv_2d(filter = 16, kernel_size = c(3,3), padding = "same") %>%
  layer_activation_leaky_relu(0.5) %>%
  layer_batch_normalization() %>%

  # Use max pooling
  layer_max_pooling_2d(pool_size = c(2,2)) %>%
  layer_dropout(0.25) %>%
  
  # Flatten max filtered output into feature vector 
  # and feed into dense layer
  layer_flatten() %>%
  layer_dense(100) %>%
  layer_activation("relu") %>%
  layer_dropout(0.5) %>%

  # Outputs from dense layer are projected onto output layer
  layer_dense(output_n) %>% 
  layer_activation("softmax")

# compile
model %>% compile(
  loss = "categorical_crossentropy",
  optimizer = optimizer_rmsprop(lr = 0.0001, decay = 1e-6),
  metrics = "accuracy"
)
```

```{r}
# fit
hist <- model %>% fit_generator(
  # training data
  train_image_array_gen,
  
  # epochs
  steps_per_epoch = as.integer(train_samples / batch_size), 
  epochs = epochs, 
  
  # validation data
  validation_data = valid_image_array_gen,
  validation_steps = as.integer(valid_samples / batch_size)
)
```

```{r}
plot(hist)
```

```{r}
test_image_files_path = "/Users/shiringlander/Documents/Github/Data/fruits-360/Test/"

test_datagen <- image_data_generator(rescale = 1/255)

test_generator <- flow_images_from_directory(
        test_image_files_path,
        test_datagen,
        target_size = target_size,
        class_mode = "categorical",
        classes = fruit_list,
        batch_size = 1,
        shuffle = FALSE,
        seed = 42)
```

```{r}
test_generator$reset()
model %>%
  evaluate_generator(test_generator, 
                     steps = as.integer(test_generator$n))
```

```{r}
classes <- test_generator$classes %>%
  factor() %>%
  table() %>%
  as.tibble()
colnames(classes)[1] <- "value"
```

```{r}
# create library of indices & class labels
indices <- train_image_array_gen$class_indices %>%
  as.data.frame() %>%
  gather() %>%
  mutate(value = as.character(value)) %>%
  left_join(classes, by = "value")
```

```{r}
# predict on test data
test_generator$reset()
predictions <- model %>% 
  predict_generator(
    generator = test_generator,
    steps = as.integer(test_generator$n)
    ) %>%
  round() %>%
  as.tibble()

colnames(predictions) <- indices$key

predictions <- predictions %>%
  mutate(truth_idx = as.character(test_generator$classes)) %>%
  left_join(indices, by = c("truth_idx" = "value"))
```

```{r}
pred_analysis <- predictions %>%
  mutate(img_id = seq(1:test_generator$n)) %>%
  gather(pred_lbl, y, Kiwi:Pomegranate) %>%
  group_by(img_id) %>%
  filter(y == max(y)) %>%
  arrange(img_id) %>%
  group_by(key, n, pred_lbl) %>%
  count()
```

```{r}
pred_analysis %>%
  mutate(percentage_pred = nn / n * 100) %>%
  ggplot(aes(x = key, y = pred_lbl, fill = percentage_pred)) +
    geom_tile()
```

```{r}
pred_analysis %>%
  mutate(prediction = case_when(
    key == pred_lbl ~ "correct",
    TRUE ~ "false"
  )) %>%
  group_by(key, prediction, n) %>%
  summarise(sum = sum(nn)) %>%
  mutate(percentage_pred = sum / n * 100) %>%
  ggplot(aes(x = key, y = prediction, fill = percentage_pred)) +
    geom_tile() +
    coord_flip()
```

---

```{r}
devtools::session_info()
```


