---
title: "Baby weight app"
draft: true
author: Shirin Elsinghorst
date: '2019-08-06'
categories: ["R"]
tags: [""]
thumbnailImagePosition: left
thumbnailImage: 
metaAlignment: center
coverMeta: out
slug: 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Libraries

```{r}
library(readr)
library(tidyverse)
library(lubridate)
library(plotly)
```

## Reference data

https://www.who.int/childgrowth/standards/weight_for_age/en/

Weight-for-age tables
percentiles

Weight-for-age: Birth to 13 weeks

```{r eval=FALSE}
boys_p_0_13 <- read_delim("/Users/shiringlander/Documents/Github/shirinsplayground/data/who_wfa/tab_wfa_boys_p_0_13.txt", delim = "\t") %>%
  mutate(gender = "boys")
girls_p_0_13 <- read_delim("/Users/shiringlander/Documents/Github/shirinsplayground/data/who_wfa/tab_wfa_girls_p_0_13.txt", delim = "\t") %>%
  mutate(gender = "girls")
p_0_13 <- bind_rows(boys_p_0_13, girls_p_0_13) %>%
  gather("percentile", "weight", P01:P999)
write_csv(p_0_13, "/Users/shiringlander/Documents/Github/who_baby_weight_app/data/p_0_13.csv")
head(p_0_13)
```

Weight-for-age: Birth to 5 years

```{r eval=FALSE}
boys_p_0_5 <- read_delim("/Users/shiringlander/Documents/Github/shirinsplayground/data/who_wfa/tab_wfa_boys_p_0_5.txt", delim = "\t") %>%
  mutate(gender = "boys")
girls_p_0_5 <- read_delim("/Users/shiringlander/Documents/Github/shirinsplayground/data/who_wfa/tab_wfa_girls_p_0_5.txt", delim = "\t") %>%
  mutate(gender = "girls")
p_0_5 <- bind_rows(boys_p_0_5, girls_p_0_5) %>%
  gather("percentile", "weight", P01:P999)
write_csv(p_0_5, "/Users/shiringlander/Documents/Github/who_baby_weight_app/data/p_0_5.csv")
head(p_0_5)
```

```{r}
p_0_5 <- read_csv("/Users/shiringlander/Documents/Github/who_baby_weight_app/data/p_0_5.csv")
p_0_13 <- read_csv("/Users/shiringlander/Documents/Github/who_baby_weight_app/data/p_0_13.csv")
```

## Adding own measurements

Needs

- `date` column
- `weight` column in kg
- first row has to contain the birthdate and birthweight

```{r, eval=FALSE}
weight_measures <- read_csv("/Users/shiringlander/Documents/Github/who_baby_weight/data/weight_measures.csv")
head(weight_measures)
```

```{r, eval=FALSE}
## Create fake data based on real values by
## adding/subtracting to/from every value a random number between -200 & 200
random_nos <- base::sample(-200:200, nrow(weight_measures), replace = TRUE)
weight_measures$weight <- weight_measures$weight + random_nos
write_csv(weight_measures, "/Users/shiringlander/Documents/Github/who_baby_weight/data/weight_measures.csv")
```

```{r}
weight_measures <- read_csv("/Users/shiringlander/Documents/Github/who_baby_weight_app/data/weight_measures.csv",
                              col_types = list(col_date(format = "%d.%m.%Y"), col_double()))
head(weight_measures)
```

```{r}
combine_measures_who <- function(weight_measures, 
                                 age_range, # "0_13" or "0_5"
                                 weight_in, # "g" or "kg"
                                 gender) { # "boy" or "girl"
  
  if (weight_in == "g") {
    weight_measures <- weight_measures %>%
      mutate(weight = weight / 1000)
  }
  
  reference_date <- weight_measures$date[[1]]
  end_date <- weight_measures$date[[nrow(weight_measures)]] + 31
  
  reference_weight <- weight_measures$weight[[1]]
  
  if (age_range == "0_13") {
    if (gender == "boy") {
      who <- p_0_13 %>%
        filter(gender == "boys")
      
    } else if (gender == "girl") {
      who <- p_0_13 %>%
        filter(gender == "girls")
      
    } else {
      stop('"gender" either has to be "boy" or "girl"')
    }
    
    who <- who %>%
    mutate(date = rep(seq(from = reference_date, by = "week", length.out = 14), 15),
         ref = "WHO")
    
    who_f <- who %>%
      filter(Week == 0)
    
  } else if (age_range == "0_5") {
    if (gender == "boy") {
      who <- p_0_5 %>%
        filter(gender == "boys")
      
    } else if (gender == "girl") {
      who <- p_0_5 %>%
        filter(gender == "girls")
      
    } else {
      stop('"gender" either has to be "boy" or "girl"')
    }
    
    who <- who %>%
      mutate(date = rep(seq(from = reference_date, by = "month", length.out = 61), 15),
         ref = "WHO") %>%
      filter(date <= !!paste(end_date))
    
    who_f <- who %>%
      filter(Month == 0)
    
  } else {
    stop('"age_range" either has to be "0_13" or "0_5"')
  }
  
  who <- weight_measures %>%
    mutate(ref = "measurement",
           percentile = "measurement",
           starting_p = who_f[which.min(abs(who_f$weight - reference_weight)), ]$percentile) %>%
    bind_rows(who)
  
  return(who)
}
```

```{r echo=FALSE, eval=FALSE}
combine_measures_who_2 <- function(weight_measures, 
                                 weight_in, # "g" or "kg"
                                 gender) { # "boy" or "girl"
  
  if (weight_in == "g") {
    weight_measures <- weight_measures %>%
      mutate(weight = weight / 1000)
  }
  
  reference_date <- weight_measures$date[[1]]
  end_date <- weight_measures$date[[nrow(weight_measures)]] + 31
  
  reference_weight <- weight_measures$weight[[1]]
  
  p_0_13_2 <- p_0_13 %>%
    mutate(date = rep(seq(from = reference_date, by = "week", length.out = 14), 30),
         ref = "WHO")
  
  p_0_5_2 <- p_0_5 %>%
      mutate(date = rep(seq(from = reference_date, by = "month", length.out = 61), 30),
         ref = "WHO")
  
  who <- p_0_13_2 %>%
    bind_rows(p_0_5_2) %>%
    filter(date <= !!paste(end_date))
  
  if (gender == "boy") {
      who <- who %>%
        filter(gender == "boys")
      
    } else if (gender == "girl") {
      who <- who %>%
        filter(gender == "girls")
      
    } else {
      stop('"gender" either has to be "boy" or "girl"')
    }
  
  who_f <- who %>%
    filter(Week == 0)
      
  who <- weight_measures %>%
    mutate(ref = "measurement",
           percentile = "measurement",
           starting_p = who_f[which.min(abs(who_f$weight - reference_weight)), ]$percentile) %>%
    bind_rows(who)
  
  return(who)
}
```

```{r}
combine_measures_who_0_13 <- combine_measures_who(weight_measures, 
                                             age_range = "0_13",
                                             weight_in = "g",
                                             gender = "boy")
```

```{r}
plot_0_13 <- combine_measures_who_0_13 %>%
  ggplot(aes(date, weight,
             linetype = ref,
             group = percentile, 
             color = percentile)) +
    geom_line() +
    geom_point() +
    xlab("Date") + ylab("Weight in kg") +
    ggtitle("Weekly WHO Percentiles for boys 0 to 13 weeks old") + 
    theme_bw()

ggplotly(plot_0_13)
```


```{r}
combine_measures_who_0_5 <- combine_measures_who(weight_measures, 
                                             age_range = "0_5",
                                             weight_in = "g",
                                             gender = "boy")

plot_0_5 <- combine_measures_who_0_5 %>%
  ggplot(aes(date, weight,
             linetype = ref,
             #group = percentile, 
             color = percentile)) +
    geom_line() +
    geom_point() +
    xlab("Date") + ylab("Weight in kg") +
    ggtitle("Monthly WHO Percentiles for boys 0 to 5 years old") + 
    theme_bw()

ggplotly(plot_0_5)
```

```{r}
# add missing dates
reference_date <- weight_measures$date[[1]]
end_date <- weight_measures$date[[nrow(weight_measures)]]

all_dates <- seq(from = reference_date, to = end_date, by = "day") %>%
  as.tibble()
colnames(all_dates) = "date"

weight_measures_all <- weight_measures %>%
  right_join(all_dates, by = "date")
```

```{r}
## approximate missing values
weight_measures_all <- weight_measures_all %>%
  mutate(weight_approx = approx(weight, n = nrow(.))[[2]])
```

```{r}
## add running week number
nos <- rep(1:ceiling(nrow(weight_measures_all)/7), each = 7)
weight_measures_all$week <- nos[1:nrow(weight_measures_all)]
```

```{r}
## calculate sum over week
weight_measures_all <- weight_measures_all %>%
  mutate(diff_day = c(0, diff(weight_approx, lag = 1)),
         diff_week = c(rep(0, 7), diff(weight_approx, lag = 7)))
```

```{r}
# https://stackoverflow.com/a/26640698
elapsed_months <- function(end_date, start_date) {
    ed <- as.POSIXlt(end_date)
    sd <- as.POSIXlt(start_date)
    mo <- (ed$year - sd$year) + (ed$mon - sd$mon)
    ifelse(day(ed) >= day(sd), mo + 1, mo)
}

weight_measures_all <- weight_measures_all %>%
  mutate(month = elapsed_months(date, reference_date))
```

https://www.stillkinder.de/gewicht-und-wachstum-von-gestillten-kindern/

Am 10.-14. Lebenstag sollte es sein Geburtsgewicht wieder erreicht haben.

Nach 3 Wochen muss es sein Geburtsgewicht wieder erreicht haben.

In den ersten 2 Monaten nimmt es normalerweise 170 – 330 g pro Woche zu.

Im 3. – 4. Monat 110 – 330 g pro Woche, 5. – 6. Monat 70 – 140 g pro Woche und vom 7. – 12. Monat 40 – 110 g pro Woche.

```{r}
weight_measures_all <- weight_measures_all %>%
  distinct(week, .keep_all = TRUE) %>%
  mutate(color = ifelse(month <= 2 & diff_week >= 170, "ok",
                        ifelse(month == 3 & diff_week >= 110, "ok",
                               ifelse(month == 4 & diff_week >= 110, "ok",
                                      ifelse(month == 5 & diff_week >= 70, "ok",
                                             ifelse(month == 6 & diff_week >= 70, "ok",
                                                    ifelse(month >=7 & diff_week >= 40, "ok", "low")))))))
```

```{r}
weight_measures_all %>%
  ggplot(aes(x = date, y = diff_week, fill = color, color = color)) +
  geom_bar(stat = "identity") +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1")
```

```{r}
## find reference percentile
reference_date <- weight_measures$date[[1]]
starting_p <- combine_measures_who_0_5 %>% 
  filter(date == reference_date) %>%
  select(starting_p) %>%
  .[[1]] %>%
  .[1]
```

```{r}
test_curves <- combine_measures_who_0_5 %>%
  filter(percentile == !!paste(starting_p)) %>%
  left_join(select(weight_measures_all, date, weight_approx), by = "date") %>%
  tail(3)
```

```{r}
cor(test_curves$weight, test_curves$weight_approx, use = "complete.obs", method = "kendall")
```

```{r, eval=FALSE, echo=FALSE}
mod <- lm(weight ~ weight_approx, test_curves)
modelr::rmse(mod, test_curves)
```

---

```{r}
sessionInfo()
```

